---
title: "Introduction to openMalariaUtilities"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to openMalariaUtilities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(openMalariaUtilities)
```

**This document is still work in progress**

Open Malaria is a powerful tool for simulating and studying the epidemiology of
malaria. However, it can also be daunting. This R package is designed to make
the use of Open Malaria easier.

## Prerequisites

Make sure that `SQLite` and `Open Malaria` are both installed and available on
the system. Please note: Even though the package should work on Windows, it is
developed and tested under Linux.

To install the package, run

```{r, eval = FALSE}
devtools::install_github("SwissTPH/r-openMalariaUtilities")
```

## Introduction

Withing this tutorial, we will create and run example scenarios. Each scenario
needs to be defined in a XML file which serves as an input for Open Malaria
(OM). OpenMalariaUtilities (OMU) contains many functions which help to generate
these input files. In this tutorial we will re-create the example from Open
Malaria itself.

## XML file generation

A large part of OMU is dedicated at the generation of valid OM XML files. Before
we start with the generation of these files, we will explain the folder layout
of your projects.

In the folder hierarchy, your working or project directory is the top level.
Within this folder, you can have one or more folders which correspond to the
names of experiments. The simulations of these experiments are controlled via
one or multiple R scripts, placed in the project directory. Please note: Most of
the directories can be modified via function arguments, but for the sake of
simplicity we using the defaults here.

```
MyProject
├── experiment_1
├── experiment_2
├── ...
├── script_1.R
├── script_2.R
└── ...
```

Each of the experiment folders will have a similar layout and content.

```
MyProject/experiment_1
├── cache/
├── logs/
├── outputs/
├── scenarios/
├── autoRegressionParameters.csv
├── densities.csv
├── base.xml
└── scenario_44.xsd
```

You do not need to worry about the creation of the folders and files. OMU will
create necessary folders and download required files for you.

The `cache` folder contains R objects which are used by OMU to store reusable
information. This will speed up certain operations. Log files are stored in the
`logs` folder (which will contain sub-folders for each step in the process) and
OM's raw output will be `outputs`. The XML files for each simulation scenario
are stored in the `scenarios` folder.

### Starting the project

In your R script, you will probably start with `library(openMalariaUtilities)`.
Thus, the beginning could look like

```{r start}
library(openMalariaUtilities)

## Basic skeleton
baseList <- list(
  ## Mandatory
  expName = "example",
  ## Mandatory
  OMVersion = 44L,
  ## Optional
  ## analysisNo = 1L,
  ## Mandatory
  demography = list(),
  monitoring = list(),
  interventions = list(),
  healthSystem = list(),
  entomology = list(),
  ## These are optional for OM
  ## parasiteGenetics = list(),
  ## pharmacology = list(),
  ## diagnostics = list(),
  model = list()
)
```

For XML file generation, OMU uses a list as input (here: `baseList`). This list
has a very [similar structure](https://swisstph.github.io/openmalaria/schema-latest.html)
as the XML file required by OM but is easier to edit from R compared
to XML. The above snippet corresponds to the minimal skeleton of this input
list. `expName` is the name of your experiment and `OMVersion` the version of
Open Malaria. Currently, only version 44 is supported. Also note, that the OM
version needs to be an integer.

Each of the six nested lists `demography`, `monitoring`, `interventions`,
`healthSystem`, `entomology` and `model` need to be filled for OM.

### Demography

There are two ways to add the data to the `demography` entry:

1. Add it manually, e.g. `baseList[["demography"]] <- list(name = "Ifakara",
   ...)`
2. Use the provided `defineDemography()` function

We will use the second option in this tutorial. You can find various `defineXY`
functions which are each designed to add a specific entry to the list. However,
you always have the option to write parts or change specific entries manually
(See also `extractList()`).

```{r demography}
demo_data <- data.frame(
  poppercent = c(
    3.474714994, 12.76004028, 14.52151394, 12.75565434, 10.83632374,
    8.393312454, 7.001421452, 5.800587654, 5.102136612, 4.182561874,
    3.339409351, 2.986112356, 2.555766582, 2.332763433, 1.77400255,
    1.008525491, 0.74167341, 0.271863401, 0.161614642
  ),
  upperbound = c(
    1, 5, 10, 15, 20, 25, 30, 35, 40, 45, 50, 55, 60, 65, 70, 75, 80, 85, 90
  )
)


baseList <- defineDemography(
  baseList,
  name = "Ifakara",
  popSize = 4000L,
  maximumAgeYrs = 90,
  lowerbound = 0,
  poppercent = demo_data$poppercent,
  upperbound = demo_data$upperbound
)

str(baseList, max.level = 2)
```

In the above snippet, we started by creating a data frame which contains the
demographic data. This in turn was used as input for `defineDemography()` among
other parameters. The function will take care that the information is added at
the correct position and also check, that the correct data types are used.

### Monitoring

The `monitoring` section controls the output of Open Malaria, which is generated
during the conducted
[surveys](https://swisstph.github.io/openmalaria/schema-latest.html).

```{r monitoring-1}
baseList[["monitoring"]] <- list(
  name = "Quarterly Surveys",
  ## Mandatory, different from OM schema
  startDate = "2000-01-01",
  continuous = monitoringContinuousGen(
    period = 1,
    options = list(
      name = c("input EIR", "simulated EIR", "human infectiousness", "N_v0",
               "immunity h", "immunity Y", "new infections",
               "num transmitting humans", "ITN coverage", "GVI coverage", "alpha",
               "P_B", "P_C*P_D"),
      value = c("true", "true", "true", "true", "true", "true", "true", "true",
                "true", "true", "true", "false", "false")
    )
  ),
  SurveyOptions = monitoringSurveyOptionsGen(
    options = list(
      name = c("nHost", "nPatent", "nUncomp", "nSevere", "nDirDeaths",
               "inputEIR", "simulatedEIR"),
      value = c("true", "true", "true", "true", "false", "true", "true")
    )
  )
)
```

We started the monitoring definition by giving it a name and defining a start
date. Here, this is mandatory because it allows us to use dates as input for
Open Malaria. It defines the starting point of the survey period.

For the options, two generator functions are used which take the name of the
option and its value as input.

```{r monitoring-2}
baseList[["monitoring"]] <- c(
  baseList[["monitoring"]],
  list(
    surveys =  monitoringSurveyTimesGen(
      detectionLimit = 40, startDate = "2000-01-01", endDate = "2020-01-01",
      interval = "quarterly", simStart = "1980-01-01"
    ),
    ageGroup = surveyAgeGroupsGen(
      lowerbound = 0,
      upperbounds = c(5, 90)
    )
  )
)

str(baseList[["monitoring"]], max.level = 1)
```

Similar to the age groups in the demography section, `surveyAgeGroupsGen()` adds
the desired age groups for the surveys.

The definition of the survey times itself is done above. Please note, that you
really should use `monitoringSurveyTimesGen()` as this function will make sure
your desired survey points are valid (and adjust them if necessary) and will
store them for later use in the cache.

We can inspect the cache to see if our survey times have been generated
correctly.

```{r monitoring-3}
listCache()
```

We can notice two things:

1. We are expecting 81 surveys, however we have 85
2. The dates have been adjusted

OMU adds one more year (here: 4 surveys) two the survey period. This is done in
order to ensure that all deployments show their effect. The dates are adjusted
so they match OM's internal time steps. For more details on this, [see
here](https://github.com/SwissTPH/openmalaria/wiki/ModelTimeUpdates#specifying-times).

Furthermore, the number of days or time steps is counted from the simulation
start `simStart`, which is here the 1st January 1980. It is recommended to start
the simulation 20 - 50 years before the first survey.

### Interventions

In this section we will define actual measures which will be used throughout the
simulation.

```{r interventions}
baseList[["interventions"]] <- list(
  name = "test",
  human = list(
    component = list(
      id = "GVI",
      name = "DDT test",
      GVI = list(
        decay = list(
          L = "0.5",
          "function" = "exponential"
        ),
        anophelesParams = list(
          mosquito = "gambiae_ss", propActive = "1",
          deterrency = list(value = "0.56"),
          preprandialKillingEffect = list(value = "0"),
          postprandialKillingEffect = list(value = "0.24")
        )
      )
    ),
    deployment = list(
      name = "DDT test",
      component = list(id = "GVI"),
      timed = list(
        deploy = list(coverage = "0.9", time = "10y")
      )
    )
  )
)
```

### Health system

```{r healtsys}
baseList[["healthSystem"]] <- list(
  ImmediateOutcomes = list(
    name = "Tanzania ACT",
    drugRegimen = list(
      firstLine = "ACT",
      inpatient = "QN",
      secondLine = "ACT"
    ),
    initialACR = list(
      ACT = list(value = 0.85),
      QN = list(value = 0.998),
      selfTreatment = list(value = 0.63)
    ),
    compliance = list(
      ACT = list(value = 0.9),
      selfTreatment = list(value = 0.85)
    ),
    nonCompliersEffective = list(
      ACT = list(value = 0),
      selfTreatment = list(value = 0)
    ),
    treatmentActions = list(
      ACT = list(
        name = "clear blood-stage infections",
        clearInfections = list(
          stage = "blood",
          timesteps = "1"
        )
      ),
      QN = list(
        name = "clear blood-stage infections",
        clearInfections = list(
          stage = "blood",
          timesteps = "1"
        )
      )
    ),
    pSeekOfficialCareUncomplicated1 = list(value = 0.04),
    pSelfTreatUncomplicated = list(value = 0.01),
    pSeekOfficialCareUncomplicated2 = list(value = 0.04),
    pSeekOfficialCareSevere = list(value = 0.48)
  ),
  CFR = list(
    group = list(lowerbound = 0, value = 0.09189),
    group = list(lowerbound = 0.25, value = 0.0810811),
    group = list(lowerbound = 0.75, value = 0.0648649),
    group = list(lowerbound = 1.5, value = 0.0689189),
    group = list(lowerbound = 2.5, value = 0.0675676),
    group = list(lowerbound = 3.5, value = 0.0297297),
    group = list(lowerbound = 4.5, value = 0.0459459),
    group = list(lowerbound = 7.5, value = 0.0945946),
    group = list(lowerbound = 12.5, value = 0.1243243),
    group = list(lowerbound = 15, value = 0.1378378)
  ),
  ## Mandatory
  pSequelaeInpatient = list(
    interpolation = "none",
    group = list(lowerbound = "0.0", value = 0.0132),
    group = list(lowerbound = "5.0", value = 0.005)
  )
)
```

### Entomology

```{r entomology}
baseList[["entomology"]] <- list(
  mode = "dynamic", name = "Namawala",
  vector = list(
    anopheles = list(
      mosquito = "gambiae_ss", propInfected = 0.078, propInfectious = "0.021",
      seasonality = list(
        annualEIR = "16", input = "EIR",
        fourierSeries = list(
          EIRRotateAngle = "0",
          coeffic = list(
            a = "0.8968", b = "2.678"
          ),
          coeffic = list(a = "-0.4551", b = "2.599")
        )
      ),
      mosq = list(
        minInfectedThreshold = "0.001",
        mosqRestDuration = list(value = "3"),
        extrinsicIncubationPeriod = list(value = "11"),
        mosqLaidEggsSameDayProportion = list(value = "0.313"),
        mosqSeekingDuration = list(value = "0.33"),
        mosqSurvivalFeedingCycleProbability = list(value = "0.623"),
        availability = list(distr = "const"),
        mosqProbBiting = list(mean = "0.95", variance = "0"),
        mosqProbFindRestSite = list(mean = "0.95", variance = "0"),
        mosqProbResting = list(mean = "0.99", variance = "0"),
        mosqProbOvipositing = list(value = "0.88"),
        mosqHumanBloodIndex = list(value = "0.939")
      ),
      nonHumanHosts = list(
        name = "unprotectedAnimals",
        mosqRelativeEntoAvailability = list(value = "1.0"),
        mosqProbBiting = list(value = "0.95"),
        mosqProbFindRestSite = list(value = "0.95"),
        mosqProbResting = list(value = "0.99")
      )
    ),
    nonHumanHosts = list(name = "unprotectedAnimals", number = "1.0")
  )
)
```

### Model

```{r model}
baseList[["model"]] <- list(
  ModelOptions = list(
    option = list(name = "LOGNORMAL_MASS_ACTION", value = "true"),
    option = list(name = "NO_PRE_ERYTHROCYTIC", value = "true"),
    option = list(name = "INNATE_MAX_DENS", value = "false"),
    option = list(name = "INDIRECT_MORTALITY_FIX", value = "false"),
    option = list(name = "NON_MALARIA_FEVERS", value = "true")
  ),
  clinical = list(
    healthSystemMemory = "6",
    NonMalariaFevers = list(
      incidence = list(
        group = list(lowerbound = "0", value = "0.322769924518357"),
        group = list(lowerbound = "0", value = "0.308520194304172"),
        group = list(lowerbound = "1", value = "0.279441774808493"),
        group = list(lowerbound = "2", value = "0.250431781111273"),
        group = list(lowerbound = "3", value = "0.223285859756841"),
        group = list(lowerbound = "4", value = "0.199298352451799"),
        group = list(lowerbound = "5", value = "0.179376872365614"),
        group = list(lowerbound = "6", value = "0.163623659390782"),
        group = list(lowerbound = "7", value = "0.152227726923469"),
        group = list(lowerbound = "8", value = "0.145022785567758"),
        group = list(lowerbound = "9", value = "0.141493087461765"),
        group = list(lowerbound = "10", value = "0.140473293219353"),
        group = list(lowerbound = "11", value = "0.141109775159515"),
        group = list(lowerbound = "12", value = "0.142644475217328"),
        group = list(lowerbound = "13", value = "0.144335079395766"),
        group = list(lowerbound = "14", value = "0.145964032924869"),
        group = list(lowerbound = "15", value = "0.147708915135714"),
        group = list(lowerbound = "16", value = "0.149731543445568"),
        group = list(lowerbound = "17", value = "0.151887428568276"),
        group = list(lowerbound = "18", value = "0.154060663485195"),
        group = list(lowerbound = "19", value = "0.156179169710494"),
        group = list(lowerbound = "20", value = "0.158135015380583"),
        group = list(lowerbound = "21", value = "0.159704766482219"),
        group = list(lowerbound = "22", value = "0.160807788387655"),
        group = list(lowerbound = "23", value = "0.161427976448279"),
        group = list(lowerbound = "24", value = "0.161620429119137"),
        group = list(lowerbound = "25", value = "0.16144021875986"),
        group = list(lowerbound = "26", value = "0.160943264630612"),
        group = list(lowerbound = "27", value = "0.160217573697398"),
        group = list(lowerbound = "28", value = "0.159422614374451"),
        group = list(lowerbound = "29", value = "0.158542519631641"),
        group = list(lowerbound = "30", value = "0.157501217628248"),
        group = list(lowerbound = "31", value = "0.156175160594841"),
        group = list(lowerbound = "32", value = "0.154402302191411"),
        group = list(lowerbound = "33", value = "0.152102040636481"),
        group = list(lowerbound = "34", value = "0.14921450014676"),
        group = list(lowerbound = "35", value = "0.145714433541659"),
        group = list(lowerbound = "36", value = "0.141800502067518"),
        group = list(lowerbound = "37", value = "0.137916853907569"),
        group = list(lowerbound = "38", value = "0.134503529382102"),
        group = list(lowerbound = "39", value = "0.131746276580642"),
        group = list(lowerbound = "40", value = "0.12969902537497"),
        group = list(lowerbound = "41", value = "0.128398077347679"),
        group = list(lowerbound = "42", value = "0.127864136551891"),
        group = list(lowerbound = "43", value = "0.12804497197004"),
        group = list(lowerbound = "44", value = "0.128894055047661"),
        group = list(lowerbound = "45", value = "0.130350838992718"),
        group = list(lowerbound = "46", value = "0.132286605622701"),
        group = list(lowerbound = "47", value = "0.134599921072495"),
        group = list(lowerbound = "48", value = "0.137212726976988"),
        group = list(lowerbound = "49", value = "0.140035253913284"),
        group = list(lowerbound = "50", value = "0.142934573453621"),
        group = list(lowerbound = "51", value = "0.145830221511879"),
        group = list(lowerbound = "52", value = "0.148674810561069"),
        group = list(lowerbound = "53", value = "0.151497963594518"),
        group = list(lowerbound = "54", value = "0.15438856687865"),
        group = list(lowerbound = "55", value = "0.157403790093505"),
        group = list(lowerbound = "56", value = "0.16059513222516"),
        group = list(lowerbound = "57", value = "0.16402433342886"),
        group = list(lowerbound = "58", value = "0.16770481415944"),
        group = list(lowerbound = "59", value = "0.171626873047865"),
        group = list(lowerbound = "60", value = "0.175748327054247"),
        group = list(lowerbound = "61", value = "0.180030857856799"),
        group = list(lowerbound = "62", value = "0.184411365583771"),
        group = list(lowerbound = "63", value = "0.188816421789366"),
        group = list(lowerbound = "64", value = "0.19316997803338"),
        group = list(lowerbound = "65", value = "0.197435603275487"),
        group = list(lowerbound = "66", value = "0.201578808813379"),
        group = list(lowerbound = "67", value = "0.205556806881398"),
        group = list(lowerbound = "68", value = "0.209307183457343"),
        group = list(lowerbound = "69", value = "0.212783260344084"),
        group = list(lowerbound = "70", value = "0.215944154621391"),
        group = list(lowerbound = "71", value = "0.218749275266548"),
        group = list(lowerbound = "72", value = "0.221187990639016"),
        group = list(lowerbound = "73", value = "0.223361260399378"),
        group = list(lowerbound = "74", value = "0.225363436789592"),
        group = list(lowerbound = "75", value = "0.227254280093211"),
        group = list(lowerbound = "76", value = "0.229084576349576"),
        group = list(lowerbound = "77", value = "0.230891971097789"),
        group = list(lowerbound = "78", value = "0.232690225166173"),
        group = list(lowerbound = "79", value = "0.234484973338876"),
        group = list(lowerbound = "80", value = "0.236276361586796"),
        group = list(lowerbound = "81", value = "0.238064394629696"),
        group = list(lowerbound = "82", value = "0.239849077182917"),
        group = list(lowerbound = "83", value = "0.241630413957381"),
        group = list(lowerbound = "84", value = "0.243408409659591"),
        group = list(lowerbound = "85", value = "0.245183068991633"),
        group = list(lowerbound = "86", value = "0.246954396651183"),
        group = list(lowerbound = "87", value = "0.248722397331501"),
        group = list(lowerbound = "88", value = "0.250487075721441"),
        group = list(lowerbound = "89", value = "0.252248436505447"),
        group = list(lowerbound = "90", value = "0.253127874257909")
      )
    )
  ),
  human = list(
    availabilityToMosquitoes = list(
      group = list(lowerbound = "0", value = "0.7076"),
      group = list(lowerbound = "0", value = "0.8538"),
      group = list(lowerbound = "5", value = "1.0"),
      group = list(lowerbound = "5", value = "1.0")
    )
  ),
  parameters = list(
    interval = "5", iseed = "1", latentp = "3",
    parameter = list(include = "false", name = "'-ln(1-Sinf)'", number = "1", value = "0.050736"),
    parameter = list(include = "false", name = "Estar", number = "2", value = "0.03247"),
    parameter = list(include = "false", name = "Simm", number = "3", value = "0.138161050830301"),
    parameter = list(include = "false", name = "Xstar_p", number = "4", value = "1514.385853233699891"),
    parameter = list(include = "false", name = "gamma_p", number = "5", value = "2.03692533424484"),
    parameter = list(include = "false", name = "sigma2i", number = "6", value = "10.173598698525799"),
    parameter = list(include = "false", name = "CumulativeYstar", number = "7", value = "35158523.31132510304451"),
    parameter = list(include = "false", name = "CumulativeHstar", number = "8", value = "97.334652723897705"),
    parameter = list(include = "false", name = "'-ln(1-alpha_m)'", number = "9", value = "2.33031045876193"),
    parameter = list(include = "false", name = "decay_m", number = "10", value = "2.53106547375805"),
    parameter = list(include = "false", name = "sigma2_0", number = "11", value = "0.655747311168152"),
    parameter = list(include = "false", name = "Xstar_v", number = "12", value = "0.916181104713054"),
    parameter = list(include = "false", name = "Ystar2", number = "13", value = "6502.26335600001039"),
    parameter = list(include = "false", name = "alpha", number = "14", value = "142601.912520000012591"),
    parameter = list(include = "false", name = "Density bias (non Garki)", number = "15", value = "0.177378570987455"),
    parameter = list(include = "false", name = "        sigma2        ", number = "16", value = "0.05"),
    parameter = list(include = "false", name = "log oddsr CF community", number = "17", value = "0.736202"),
    parameter = list(include = "false", name = "Indirect risk cofactor", number = "18", value = "0.018777338"),
    parameter = list(include = "false", name = "Non-malaria infant mortality", number = "19", value = "49.539046599999999"),
    parameter = list(include = "false", name = "Density bias (Garki)", number = "20", value = "4.79610772546704"),
    parameter = list(include = "false", name = "Severe Malaria Threshhold", number = "21", value = "784455.599999999976717"),
    parameter = list(include = "false", name = "Immunity Penalty", number = "22", value = "1"),
    parameter = list(include = "false", name = "Immune effector decay", number = "23", value = "0"),
    parameter = list(include = "false", name = "comorbidity intercept", number = "24", value = "0.0968"),
    parameter = list(include = "false", name = "Ystar half life", number = "25", value = "0.275437402"),
    parameter = list(include = "false", name = "Ystar1", number = "26", value = "0.596539864"),
    parameter = list(include = "false", name = "Asexual immunity decay", number = "27", value = "0"),
    parameter = list(include = "false", name = "Ystar0", number = "28", value = "296.302437899999973"),
    parameter = list(include = "false", name = "Idete multiplier", number = "29", value = "2.797523626"),
    parameter = list(include = "false", name = "critical age for comorbidity", number = "30", value = "0.117383")
  )
)
```

## Prepare to run the simulations

```{r setup-1}
## Setup dirs
setupDirs(experimentName = "exp_test", replace = TRUE)

## Copy necessary Open Malaria files.
setupOM()
```

`setupDirs()` creates the experiment folder given as an argument in the project
directory. This also includes all the sub-folders which were mentioned in the
[introduction](#introduction) .

We can verify this by checking the layout of the experiment directory.

```{r setup-2}
list.dirs("exp_test")
```

`setupOM()` will make sure that all required files for Open Malaria to run are
in place. This includes the correct `.xsd` schema file (according to the
specified OM version) as well as further supplementary files.

```{r setup-3}
list.files("exp_test")
```

## Create XML file

As the name implies, `createBaseXml()` will translate our input list to a XML
document and write it to disk.

```{r setup-4}
## Create base XML file
createBaseXml(baseList, replace = TRUE)
readLines(file("exp_test/exp_test_base.xml", "r"), n = 10)
```

## Run simulations

In order to run the simulation, we can simply call `runSimulations()`.

```{r sim-1, eval = FALSE}
## Run simulations
runSimulations()
#> [1] "Running scenario [1/1]"
```

By default, the output in the R console is fairly minimal. More detailed logs
can be found in the `logs/simulation` folder.

```{r sim-2, eval = FALSE}
list.files("exp_test/logs/simulation")
#> [1] "exp_test_base.log"
```

Open Malaria generates two [output
files](https://github.com/SwissTPH/openmalaria/wiki/UserGuide#interpretation-of-the-test-scenario)
which are both placed in the `outputs` folder:

1. The `*_cts.txt` file

```{r sim-3, eval = FALSE}
read.table("exp_test/outputs/exp_test_base_cts.txt", header = TRUE, sep = "\t", skip = 1, nrows = 10)
#>    timestep  input.EIR simulated.EIR human.infectiousness N_v0.gambiae_ss.
#> 1         0 0.00995606    0.00995049            0.0179202          30147.4
#> 2         1 0.01958060    0.15390600            0.0175667          39181.0
#> 3         2 0.03850920    0.24658900            0.0171114          46368.5
#> 4         3 0.07460740    0.37191900            0.0167741          49838.5
#> 5         4 0.14030700    0.52390000            0.0164523          48641.0
#> 6         5 0.25256000    0.68747000            0.0165751          43197.7
#> 7         6 0.42953400    0.84148200            0.0188351          35064.0
#> 8         7 0.68226100    0.96738200            0.0231314          26185.7
#> 9         8 1.00214000    1.07333000            0.0296558          18146.4
#> 10        9 1.35047000    1.18202000            0.0377866          11790.1
#>    immunity.h immunity.Y new.infections num.transmitting.humans ITN.coverage
#> 1     256.846   11899700              0                    2613            0
#> 2     256.817   11900500            400                    2590            0
#> 3     256.856   11895200            692                    2572            0
#> 4     256.934   11896300           1054                    2549            0
#> 5     257.109   11897600           1469                    2526            0
#> 6     257.299   11899200           1795                    2507            0
#> 7     257.791   11901800           2250                    2636            0
#> 8     258.367   11903800           2643                    2798            0
#> 9     259.013   11904600           2987                    3019            0
#> 10    259.681   11909900           3256                    3280            0
#>    GVI.coverage alpha_i.gambiae_ss.
#> 1             0         0.000206347
#> 2             0         0.000206342
#> 3             0         0.000206349
#> 4             0         0.000206332
#> 5             0         0.000206344
#> 6             0         0.000206344
#> 7             0         0.000206347
#> 8             0         0.000206338
#> 9             0         0.000206330
#> 10            0         0.000206329
```

2. The `*_out.txt` file

```{r sim-4, eval = FALSE}
read.table("exp_test/outputs/exp_test_base_out.txt", header = FALSE, sep = "\t", nrows = 10)
#>    V1 V2 V3          V4
#> 1   1  1  0 6.07000e+02
#> 2   1  2  0 3.39300e+03
#> 3   1  1  3 1.37000e+02
#> 4   1  2  3 1.15500e+03
#> 5   1  1 14 3.35770e+04
#> 6   1  2 14 2.94240e+04
#> 7   1  1 15 7.11000e+02
#> 8   1  2 15 4.80000e+01
#> 9   1  0 35 2.19321e-01
#> 10  1  0 36 2.87284e-01
```

The `*_out.txt` file is usually the file you are interested in, as it contains
the measurement data for each survey you defined in the input list. [See
here](https://github.com/SwissTPH/openmalaria/wiki/MonitoringOutput) for more information.

## Read results

It can be a bit difficult to interpret Open Malaria's output files, thus OMU
provides a helper function which reads an output file and assigns column names
and translates certain values (e.g. the survey dates).

```{r results-1, eval = FALSE}
outfile <- readOutputFile("exp_test/outputs/exp_test_base_out.txt")
head(outfile)
#>    survey_date third_dimension measure value
#> 1:  1999-12-27               1   nHost   607
#> 2:  1999-12-27               2   nHost  3393
#> 3:  1999-12-27               1 nPatent   137
#> 4:  1999-12-27               2 nPatent  1155
#> 5:  1999-12-27               1 nUncomp 33577
#> 6:  1999-12-27               2 nUncomp 29424
```
