---
title: "Introduction to openMalariaUtilities"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to openMalariaUtilities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(openMalariaUtilities)
```

Open Malaria is a powerful tool for simulating and studying the epidemiology of
malaria. However, it can also be daunting. This R package is designed to make
the use of Open Malaria easier.

## Prerequisites

Make sure that `SQLite` and `Open Malaria` are both installed and available on
the system. Please note: Even though the package should work on Windows, it is
developed and tested under Linux.

To install the package, run

```{r, eval = FALSE}
devtools::install_github("SwissTPH/r-openMalariaUtilities")
```

## Introduction

Withing this tutorial, we will create and run example scenarios. Each scenario
needs to be defined in a XML file which serves as an input for Open Malaria
(OM). OpenMalariaUtilities (OMU) contains many functions which help to generate
these input files.
