---
title: "Introduction to openMalariaUtilities"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to openMalariaUtilities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(openMalariaUtilities)
```

**This document is still work in progress**

Open Malaria is a powerful tool for simulating and studying the epidemiology of
malaria. However, it can also be daunting. This R package is designed to make
the use of Open Malaria easier.

## Prerequisites

Make sure that `SQLite` and `Open Malaria` are both installed and available on
the system. Please note: Even though the package should work on Windows, it is
developed and tested under Linux.

To install the package, run

```{r, eval = FALSE}
devtools::install_github("SwissTPH/r-openMalariaUtilities")
```

## Introduction

Withing this tutorial, we will create and run example scenarios. Each scenario
needs to be defined in a XML file which serves as an input for Open Malaria
(OM). OpenMalariaUtilities (OMU) contains many functions which help to generate
these input files. In this tutorial we will re-create the example from Open
Malaria itself.

## XML file generation

A large part of OMU is dedicated at the generation of valid OM XML files. Before
we start with the generation of these files, we will explain the folder layout
of your projects.

In the folder hierarchy, your working or project directory is the top level.
Within this folder, you can have one or more folders which correspond to the
names of experiments. The simulations of these experiments are controlled via
one or multiple R scripts, placed in the project directory. Please note: Most of
the directories can be modified via function arguments, but for the sake of
simplicity we using the defaults here.

```
MyProject
├── experiment_1
├── experiment_2
├── ...
├── script_1.R
├── script_2.R
└── ...
```

Each of the experiment folders will have a similar layout and content.

```
MyProject/experiment_1
├── cache/
├── logs/
├── outputs/
├── scenarios/
├── autoRegressionParameters.csv
├── densities.csv
├── base.xml
└── scenario_44.xsd
```

You do not need to worry about the creation of the folders and files. OMU will
create necessary folders and download required files for you.

The `cache` folder contains R objects which are used by OMU to store reusable
information. This will speed up certain operations. Log files are stored in the
`logs` folder (which will contain sub-folders for each step in the process) and
OM's raw output will be `outputs`. The XML files for each simulation scenario
are stored in the `scenarios` folder.

### Starting the project

In your R script, you will probably start with `library(openMalariaUtilities)`.
Thus, the beginning could look like

```{r start}
library(openMalariaUtilities)

## Basic skeleton
baseList <- list(
  ## Mandatory
  expName = "example",
  ## Mandatory
  OMVersion = 44L,
  ## Optional
  ## analysisNo = 1L,
  ## Mandatory
  demography = list(),
  monitoring = list(),
  interventions = list(),
  healthSystem = list(),
  entomology = list(),
  ## These are optional for OM
  ## parasiteGenetics = list(),
  ## pharmacology = list(),
  ## diagnostics = list(),
  model = list()
)
```

For XML file generation, OMU uses a list as input (here: `baseList`). This list
has a very [similar structure](https://swisstph.github.io/openmalaria/schema-latest.html)
as the XML file required by OM but is easier to edit from R compared
to XML. The above snippet corresponds to the minimal skeleton of this input
list. `expName` is the name of your experiment and `OMVersion` the version of
Open Malaria. Currently, only version 44 is supported. Also note, that the OM
version needs to be an integer.

Each of the six nested lists `demography`, `monitoring`, `interventions`,
`healthSystem`, `entomology` and `model` need to be filled for OM.

### Demography

There are two ways to add the data to the `demography` entry:

1. Add it manually, e.g. `baseList[["demography"]] <- list(name = "Ifakara",
   ...)`
2. Use the provided `defineDemography()` function

We will use the second option in this tutorial. You can find various `defineXY`
functions which are each designed to add a specific entry to the list. However,
you always have the option to write parts or change specific entries manually
(See also `extractList()`).

```{r demography}
demo_data <- data.frame(
  poppercent = c(
    3.474714994, 12.76004028, 14.52151394, 12.75565434, 10.83632374,
    8.393312454, 7.001421452, 5.800587654, 5.102136612, 4.182561874,
    3.339409351, 2.986112356, 2.555766582, 2.332763433, 1.77400255,
    1.008525491, 0.74167341, 0.271863401, 0.161614642
  ),
  upperbound = c(
    1, 5, 10, 15, 20, 25, 30, 35, 40, 45, 50, 55, 60, 65, 70, 75, 80, 85, 90
  )
)


baseList <- defineDemography(
  baseList,
  name = "Ifakara",
  popSize = 4000L,
  maximumAgeYrs = 90,
  lowerbound = 0,
  poppercent = demo_data$poppercent,
  upperbound = demo_data$upperbound
)

str(baseList, max.level = 2)
```

In the above snippet, we started by creating a data frame which contains the
demographic data. This in turn was used as input for `defineDemography()` among
other parameters. The function will take care that the information is added at
the correct position and also check, that the correct data types are used.

### Monitoring

The `monitoring` section controls the output of Open Malaria, which is generated
during the conducted
[surveys](https://swisstph.github.io/openmalaria/schema-latest.html).

```{r monitoring-1}
baseList[["monitoring"]] <- list(
  name = "Quarterly Surveys",
  ## Mandatory, different from OM schema
  startDate = "2000-01-01",
  continuous = monitoringContinuousGen(
    period = 1,
    options = list(
      name = c("input EIR", "simulated EIR", "human infectiousness", "N_v0",
               "immunity h", "immunity Y", "new infections",
               "num transmitting humans", "ITN coverage", "GVI coverage", "alpha",
               "P_B", "P_C*P_D"),
      value = c("true", "true", "true", "true", "true", "true", "true", "true",
                "true", "true", "true", "false", "false")
    )
  ),
  SurveyOptions = monitoringSurveyOptionsGen(
    options = list(
      name = c("nHost", "nPatent", "nUncomp", "nSevere", "nDirDeaths",
               "inputEIR", "simulatedEIR"),
      value = c("true", "true", "true", "true", "false", "true", "true")
    )
  )
)
```

We started the monitoring definition by giving it a name and defining a start
date. Here, this is mandatory because it allows us to use dates as input for
Open Malaria. It defines the starting point of the survey period.

For the options, two generator functions are used which take the name of the
option and its value as input.

```{r monitoring-2}
baseList[["monitoring"]] <- c(
  baseList[["monitoring"]],
  list(
    surveys =  monitoringSurveyTimesGen(
      detectionLimit = 40, startDate = "2000-01-01", endDate = "2020-01-01",
      interval = "quarterly", simStart = "1980-01-01"
    ),
    ageGroup = surveyAgeGroupsGen(
      lowerbound = 0,
      upperbounds = c(5, 90)
    )
  )
)

str(baseList[["monitoring"]], max.level = 1)
```

Similar to the age groups in the demography section, `surveyAgeGroupsGen()` adds
the desired age groups for the surveys.

The definition of the survey times itself is done above. Please note, that you
really should use `monitoringSurveyTimesGen()` as this function will make sure
your desired survey points are valid (and adjust them if necessary) and will
store them for later use in the cache.

We can inspect the cache to see if our survey times have been generated
correctly.

```{r monitoring-3}
listCache()
```

We can notice two things:

1. We are expecting 81 surveys, however we have 85
2. The dates have been adjusted

OMU adds one more year (here: 4 surveys) two the survey period. This is done in
order to ensure that all deployments show their effect. The dates are adjusted
so they match OM's internal time steps. For more details on this, [see
here](https://github.com/SwissTPH/openmalaria/wiki/ModelTimeUpdates#specifying-times).

Furthermore, the number of days or time steps is counted from the simulation
start `simStart`, which is here the 1st January 1980. It is recommended to start
the simulation 20 - 50 years before the first survey.

### Interventions

In this section we will define actual measures which will be used throughout the
simulation.

```{r interventions}
baseList[["interventions"]] <- list(
  name = "test",
  human = list(
    component = list(
      id = "GVI",
      name = "DDT test",
      GVI = list(
        decay = list(
          L = "0.5",
          "function" = "exponential"
        ),
        anophelesParams = list(
          mosquito = "gambiae_ss", propActive = "1",
          deterrency = list(value = "0.56"),
          preprandialKillingEffect = list(value = "0"),
          postprandialKillingEffect = list(value = "0.24")
        )
      )
    ),
    deployment = list(
      name = "DDT test",
      component = list(id = "GVI"),
      timed = list(
        deploy = list(coverage = "0.9", time = "10y")
      )
    )
  )
)
```

### Health system

```{r healtsys}
baseList[["healthSystem"]] <- list(
  ImmediateOutcomes = list(
    name = "Tanzania ACT",
    drugRegimen = list(
      firstLine = "ACT",
      inpatient = "QN",
      secondLine = "ACT"
    ),
    initialACR = list(
      ACT = list(value = 0.85),
      QN = list(value = 0.998),
      selfTreatment = list(value = 0.63)
    ),
    compliance = list(
      ACT = list(value = 0.9),
      selfTreatment = list(value = 0.85)
    ),
    nonCompliersEffective = list(
      ACT = list(value = 0),
      selfTreatment = list(value = 0)
    ),
    treatmentActions = list(
      ACT = list(
        name = "clear blood-stage infections",
        clearInfections = list(
          stage = "blood",
          timesteps = "1"
        )
      ),
      QN = list(
        name = "clear blood-stage infections",
        clearInfections = list(
          stage = "blood",
          timesteps = "1"
        )
      )
    ),
    pSeekOfficialCareUncomplicated1 = list(value = 0.04),
    pSelfTreatUncomplicated = list(value = 0.01),
    pSeekOfficialCareUncomplicated2 = list(value = 0.04),
    pSeekOfficialCareSevere = list(value = 0.48)
  ),
  CFR = list(
    group = list(lowerbound = 0, value = 0.09189),
    group = list(lowerbound = 0.25, value = 0.0810811),
    group = list(lowerbound = 0.75, value = 0.0648649),
    group = list(lowerbound = 1.5, value = 0.0689189),
    group = list(lowerbound = 2.5, value = 0.0675676),
    group = list(lowerbound = 3.5, value = 0.0297297),
    group = list(lowerbound = 4.5, value = 0.0459459),
    group = list(lowerbound = 7.5, value = 0.0945946),
    group = list(lowerbound = 12.5, value = 0.1243243),
    group = list(lowerbound = 15, value = 0.1378378)
  ),
  ## Mandatory
  pSequelaeInpatient = list(
    interpolation = "none",
    group = list(lowerbound = "0.0", value = 0.0132),
    group = list(lowerbound = "5.0", value = 0.005)
  )
)
```

### Entomology

```{r entomology}
baseList[["entomology"]] <- list(
  mode = "dynamic", name = "Namawala",
  vector = list(
    anopheles = list(
      mosquito = "gambiae_ss", propInfected = 0.078, propInfectious = "0.021",
      seasonality = list(
        annualEIR = "16", input = "EIR",
        fourierSeries = list(
          EIRRotateAngle = "0",
          coeffic = list(
            a = "0.8968", b = "2.678"
          ),
          coeffic = list(a = "-0.4551", b = "2.599")
        )
      ),
      mosq = list(
        minInfectedThreshold = "0.001",
        mosqRestDuration = list(value = "3"),
        extrinsicIncubationPeriod = list(value = "11"),
        mosqLaidEggsSameDayProportion = list(value = "0.313"),
        mosqSeekingDuration = list(value = "0.33"),
        mosqSurvivalFeedingCycleProbability = list(value = "0.623"),
        availability = list(distr = "const"),
        mosqProbBiting = list(mean = "0.95", variance = "0"),
        mosqProbFindRestSite = list(mean = "0.95", variance = "0"),
        mosqProbResting = list(mean = "0.99", variance = "0"),
        mosqProbOvipositing = list(value = "0.88"),
        mosqHumanBloodIndex = list(value = "0.939")
      ),
      nonHumanHosts = list(
        name = "unprotectedAnimals",
        mosqRelativeEntoAvailability = list(value = "1.0"),
        mosqProbBiting = list(value = "0.95"),
        mosqProbFindRestSite = list(value = "0.95"),
        mosqProbResting = list(value = "0.99")
      )
    ),
    nonHumanHosts = list(name = "unprotectedAnimals", number = "1.0")
  )
)
```

### Model

```{r model}
baseList[["model"]] <- list(
  ModelOptions = list(
    option = list(name = "LOGNORMAL_MASS_ACTION", value = "true"),
    option = list(name = "NO_PRE_ERYTHROCYTIC", value = "true"),
    option = list(name = "INNATE_MAX_DENS", value = "false"),
    option = list(name = "INDIRECT_MORTALITY_FIX", value = "false"),
    option = list(name = "NON_MALARIA_FEVERS", value = "true")
  ),
  clinical = list(
    healthSystemMemory = "6",
    NonMalariaFevers = list(
      incidence = list(
        group = list(lowerbound = "0", value = "0.322769924518357"),
        group = list(lowerbound = "0", value = "0.308520194304172"),
        group = list(lowerbound = "1", value = "0.279441774808493"),
        group = list(lowerbound = "2", value = "0.250431781111273"),
        group = list(lowerbound = "3", value = "0.223285859756841"),
        group = list(lowerbound = "4", value = "0.199298352451799"),
        group = list(lowerbound = "5", value = "0.179376872365614"),
        group = list(lowerbound = "6", value = "0.163623659390782"),
        group = list(lowerbound = "7", value = "0.152227726923469"),
        group = list(lowerbound = "8", value = "0.145022785567758"),
        group = list(lowerbound = "9", value = "0.141493087461765"),
        group = list(lowerbound = "10", value = "0.140473293219353"),
        group = list(lowerbound = "11", value = "0.141109775159515"),
        group = list(lowerbound = "12", value = "0.142644475217328"),
        group = list(lowerbound = "13", value = "0.144335079395766"),
        group = list(lowerbound = "14", value = "0.145964032924869"),
        group = list(lowerbound = "15", value = "0.147708915135714"),
        group = list(lowerbound = "16", value = "0.149731543445568"),
        group = list(lowerbound = "17", value = "0.151887428568276"),
        group = list(lowerbound = "18", value = "0.154060663485195"),
        group = list(lowerbound = "19", value = "0.156179169710494"),
        group = list(lowerbound = "20", value = "0.158135015380583"),
        group = list(lowerbound = "21", value = "0.159704766482219"),
        group = list(lowerbound = "22", value = "0.160807788387655"),
        group = list(lowerbound = "23", value = "0.161427976448279"),
        group = list(lowerbound = "24", value = "0.161620429119137"),
        group = list(lowerbound = "25", value = "0.16144021875986"),
        group = list(lowerbound = "26", value = "0.160943264630612"),
        group = list(lowerbound = "27", value = "0.160217573697398"),
        group = list(lowerbound = "28", value = "0.159422614374451"),
        group = list(lowerbound = "29", value = "0.158542519631641"),
        group = list(lowerbound = "30", value = "0.157501217628248"),
        group = list(lowerbound = "31", value = "0.156175160594841"),
        group = list(lowerbound = "32", value = "0.154402302191411"),
        group = list(lowerbound = "33", value = "0.152102040636481"),
        group = list(lowerbound = "34", value = "0.14921450014676"),
        group = list(lowerbound = "35", value = "0.145714433541659"),
        group = list(lowerbound = "36", value = "0.141800502067518"),
        group = list(lowerbound = "37", value = "0.137916853907569"),
        group = list(lowerbound = "38", value = "0.134503529382102"),
        group = list(lowerbound = "39", value = "0.131746276580642"),
        group = list(lowerbound = "40", value = "0.12969902537497"),
        group = list(lowerbound = "41", value = "0.128398077347679"),
        group = list(lowerbound = "42", value = "0.127864136551891"),
        group = list(lowerbound = "43", value = "0.12804497197004"),
        group = list(lowerbound = "44", value = "0.128894055047661"),
        group = list(lowerbound = "45", value = "0.130350838992718"),
        group = list(lowerbound = "46", value = "0.132286605622701"),
        group = list(lowerbound = "47", value = "0.134599921072495"),
        group = list(lowerbound = "48", value = "0.137212726976988"),
        group = list(lowerbound = "49", value = "0.140035253913284"),
        group = list(lowerbound = "50", value = "0.142934573453621"),
        group = list(lowerbound = "51", value = "0.145830221511879"),
        group = list(lowerbound = "52", value = "0.148674810561069"),
        group = list(lowerbound = "53", value = "0.151497963594518"),
        group = list(lowerbound = "54", value = "0.15438856687865"),
        group = list(lowerbound = "55", value = "0.157403790093505"),
        group = list(lowerbound = "56", value = "0.16059513222516"),
        group = list(lowerbound = "57", value = "0.16402433342886"),
        group = list(lowerbound = "58", value = "0.16770481415944"),
        group = list(lowerbound = "59", value = "0.171626873047865"),
        group = list(lowerbound = "60", value = "0.175748327054247"),
        group = list(lowerbound = "61", value = "0.180030857856799"),
        group = list(lowerbound = "62", value = "0.184411365583771"),
        group = list(lowerbound = "63", value = "0.188816421789366"),
        group = list(lowerbound = "64", value = "0.19316997803338"),
        group = list(lowerbound = "65", value = "0.197435603275487"),
        group = list(lowerbound = "66", value = "0.201578808813379"),
        group = list(lowerbound = "67", value = "0.205556806881398"),
        group = list(lowerbound = "68", value = "0.209307183457343"),
        group = list(lowerbound = "69", value = "0.212783260344084"),
        group = list(lowerbound = "70", value = "0.215944154621391"),
        group = list(lowerbound = "71", value = "0.218749275266548"),
        group = list(lowerbound = "72", value = "0.221187990639016"),
        group = list(lowerbound = "73", value = "0.223361260399378"),
        group = list(lowerbound = "74", value = "0.225363436789592"),
        group = list(lowerbound = "75", value = "0.227254280093211"),
        group = list(lowerbound = "76", value = "0.229084576349576"),
        group = list(lowerbound = "77", value = "0.230891971097789"),
        group = list(lowerbound = "78", value = "0.232690225166173"),
        group = list(lowerbound = "79", value = "0.234484973338876"),
        group = list(lowerbound = "80", value = "0.236276361586796"),
        group = list(lowerbound = "81", value = "0.238064394629696"),
        group = list(lowerbound = "82", value = "0.239849077182917"),
        group = list(lowerbound = "83", value = "0.241630413957381"),
        group = list(lowerbound = "84", value = "0.243408409659591"),
        group = list(lowerbound = "85", value = "0.245183068991633"),
        group = list(lowerbound = "86", value = "0.246954396651183"),
        group = list(lowerbound = "87", value = "0.248722397331501"),
        group = list(lowerbound = "88", value = "0.250487075721441"),
        group = list(lowerbound = "89", value = "0.252248436505447"),
        group = list(lowerbound = "90", value = "0.253127874257909")
      )
    )
  ),
  human = list(
    availabilityToMosquitoes = list(
      group = list(lowerbound = "0", value = "0.7076"),
      group = list(lowerbound = "0", value = "0.8538"),
      group = list(lowerbound = "5", value = "1.0"),
      group = list(lowerbound = "5", value = "1.0")
    )
  ),
  parameters = list(
    interval = "5", iseed = "1", latentp = "3",
    parameter = list(include = "false", name = "'-ln(1-Sinf)'", number = "1", value = "0.050736"),
    parameter = list(include = "false", name = "Estar", number = "2", value = "0.03247"),
    parameter = list(include = "false", name = "Simm", number = "3", value = "0.138161050830301"),
    parameter = list(include = "false", name = "Xstar_p", number = "4", value = "1514.385853233699891"),
    parameter = list(include = "false", name = "gamma_p", number = "5", value = "2.03692533424484"),
    parameter = list(include = "false", name = "sigma2i", number = "6", value = "10.173598698525799"),
    parameter = list(include = "false", name = "CumulativeYstar", number = "7", value = "35158523.31132510304451"),
    parameter = list(include = "false", name = "CumulativeHstar", number = "8", value = "97.334652723897705"),
    parameter = list(include = "false", name = "'-ln(1-alpha_m)'", number = "9", value = "2.33031045876193"),
    parameter = list(include = "false", name = "decay_m", number = "10", value = "2.53106547375805"),
    parameter = list(include = "false", name = "sigma2_0", number = "11", value = "0.655747311168152"),
    parameter = list(include = "false", name = "Xstar_v", number = "12", value = "0.916181104713054"),
    parameter = list(include = "false", name = "Ystar2", number = "13", value = "6502.26335600001039"),
    parameter = list(include = "false", name = "alpha", number = "14", value = "142601.912520000012591"),
    parameter = list(include = "false", name = "Density bias (non Garki)", number = "15", value = "0.177378570987455"),
    parameter = list(include = "false", name = "        sigma2        ", number = "16", value = "0.05"),
    parameter = list(include = "false", name = "log oddsr CF community", number = "17", value = "0.736202"),
    parameter = list(include = "false", name = "Indirect risk cofactor", number = "18", value = "0.018777338"),
    parameter = list(include = "false", name = "Non-malaria infant mortality", number = "19", value = "49.539046599999999"),
    parameter = list(include = "false", name = "Density bias (Garki)", number = "20", value = "4.79610772546704"),
    parameter = list(include = "false", name = "Severe Malaria Threshhold", number = "21", value = "784455.599999999976717"),
    parameter = list(include = "false", name = "Immunity Penalty", number = "22", value = "1"),
    parameter = list(include = "false", name = "Immune effector decay", number = "23", value = "0"),
    parameter = list(include = "false", name = "comorbidity intercept", number = "24", value = "0.0968"),
    parameter = list(include = "false", name = "Ystar half life", number = "25", value = "0.275437402"),
    parameter = list(include = "false", name = "Ystar1", number = "26", value = "0.596539864"),
    parameter = list(include = "false", name = "Asexual immunity decay", number = "27", value = "0"),
    parameter = list(include = "false", name = "Ystar0", number = "28", value = "296.302437899999973"),
    parameter = list(include = "false", name = "Idete multiplier", number = "29", value = "2.797523626"),
    parameter = list(include = "false", name = "critical age for comorbidity", number = "30", value = "0.117383")
  )
)
```

## Prepare to run the simulations

```{r setup-1}
## Setup dirs
setupDirs(experimentName = "exp_test", replace = TRUE)

## Copy necessary Open Malaria files.
setupOM()
```

## Create XML file

```{r setup-2}
## Create base XML file
createBaseXml(baseList, replace = TRUE)
```

## Run simulations

```{r sim-1, eval = FALSE}
## Run simulations
runSimulations()
```

## Collect results

```{r results-1, eval = FALSE}
readResults(getCache("experimentDir"), dbName = "test")
```
